---
title: "FirstProject"
author: "Y. Lu"
date: "2018/12/31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Obtain initial data
File is firstly unzipped and then csv file is read into a data frame.


```{r obtaining_data}

fileName="activity.zip"
unzip(fileName)
DataA=read.csv("activity.csv")
summary(DataA)
```
Then date fomat is converted to POSIXct class and interval is converted to hour:minute format.

```{r convert_data}
library(lubridate)
Sys.setlocale("LC_TIME", "English")
DataA$Time<- as.POSIXct(DataA$date, format="%Y-%m-%d")
DataA$day <- weekdays(as.Date(DataA$date))
str(DataA)
```

## Obtain the mean total step number of steps per day

#### 1. Calculate the total number of steps taken per day (ignore the missing values) and report the mean and median of the total number of steps taken per day
```{r total_steps,echo=TRUE}
Sys.setlocale("LC_TIME", "English")
DataSum <- aggregate(DataA$steps ~ DataA$date, FUN=sum)

```

#### 2. Histogram of steps on each day
```{r histo_total_steps, echo=TRUE}
hist(DataSum$`DataA$steps`, breaks=5, xlab="Steps", main = "Total Steps per Day")

```
#### 3. Calculate the mean and median of total number of steps per day

```{r mean_total_steps, echo=TRUE}
mean_steps <- mean(DataSum$steps,na.rm=TRUE)
median_steps <- median(DataSum$steps,na.rm=TRUE)
```
Mean and median of the total number of steps taken per day are `r mean_steps` steps and `r median_steps` steps.






## Average daily activity pattern

#### 1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r daily_pattern, echo=TRUE}
library(plyr,quietly=TRUE)
library(ggplot2)

cleanData <- DataA[!is.na(DataA$steps),]

intervalTable <- ddply(cleanData, .(interval), summarize, Avg = mean(steps))

p <- ggplot(intervalTable, aes(x=interval, y=Avg), xlab = "Interval", ylab="Average Number of Steps")
p + geom_line()+xlab("Interval")+ylab("Average Number of Steps")+ggtitle("Average Number of Steps per Interval")
```

#### 2. Obtain the interval with maximum steps
```{r max_interval, echo=TRUE}

maxSteps <- max(intervalTable$Avg)
max_interval<-intervalTable[intervalTable$Avg==maxSteps,1]



```
Interval **"`r max_interval`"** contains on average the maximum number of steps (**`r maxSteps` steps**).



## Imputing missing values

#### 1. report the number of missing values
```{r missing_number, echo=TRUE}
na_number <- sum(is.na(DataA$steps))

```
Total number of missing values in the dataset amounts to **`r na_number` **.

#### 2. Fill in the missing values
Hereby, we impute missing values based on average steps based on the day of the week.


```{r na_imputing, echo=TRUE}
clean <- DataA[!is.na(DataA$steps),]
avgTable <- ddply(clean, .(interval, day), summarize, Avg = mean(steps))

nadata<- DataA[is.na(DataA$steps),]

newdata<-merge(nadata, avgTable, by=c("interval", "day"))

newdata2<- newdata[,c(6,4,1,5,2)]
colnames(newdata2)<- c("steps", "date", "interval", "Time", "day")

mergeData <- rbind(clean, newdata2)

DataSum2 <- aggregate(mergeData$steps ~ mergeData$date, FUN=sum)
```


#### 3. Make a histogram of total number of steps and calculate mean and median total number

```{r histo_imputing}
colnames(DataSum2)<- c("Date", "Steps")

mean_Steps<-as.integer(mean(DataSum2$Steps))

median_Steps<-as.integer(median(DataSum2$Steps))

hist(DataSum2$Steps, breaks=5, xlab="Steps", main = "Total Steps per Day with NAs Fixed", col="Black")
hist(DataSum$`DataA$steps`, breaks=5, xlab="Steps", main = "Total Steps per Day with NAs Fixed", col="Grey", add=T)
legend("topright", c("Imputed Data", "Non-NA Data"), fill=c("black", "grey") )

```
Then mean and median steps are separately **"`r mean_Steps`"** and **"`r median_Steps`"**


## Are there differences in activity patterns between weekdays and weekends?

#### 1. creating a new factor variable in the dataset with two levels¨C ¡°weekday¡± and ¡°weekend¡± indicating whether a given date is a weekday or weekend day.
```{r factor, echo=TRUE}
mergeData$DayCategory <- ifelse(mergeData$day %in% c("Saturday", "Sunday"), "Weekend", "Weekday")

```


#### 2. Make a panel plot

```{r panel, echo=TRUE}

library(lattice) 

intervalTable2 <- ddply(mergeData, .(interval, DayCategory), summarize, Avg = mean(steps))

xyplot(Avg~interval|DayCategory, data=intervalTable2, type="l",  layout = c(1,2),
       main="Average Steps per Interval Based on Type of Day", 
       ylab="Average Number of Steps", xlab="Interval")


```

